---
title: " Assignment 3 Fundamentals of ML"
output: pdf_document
---

```{r setup, include=FALSE,message =FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

library(reshape)
library(caret)
library(e1071)


# readin the excel data into dataframe
```{r}
rm(list=ls())
getwd()
setwd("/Users/ankithdasu/Desktop/Spring 2022/Fundamentals of Machine Learning/Assignment 3")
NvBay3 <- read.csv("UniversalBank.csv")
head(NvBay3)
tail(NvBay3)
```
# Converting data into factors(categorical) mainly the one which are important to this.

```{r}
NvBay3$Personal.Loan = as.factor(NvBay3$Personal.Loan) # converting Personal Loan into categorical data
NvBay3$Online = as.factor(NvBay3$Online) # converting Online into categorical data
NvBay3$CreditCard = as.factor(NvBay3$CreditCard) # converting CreditCard into categorical data
```

#Data partition 60 % training and 40 % into validation
```{r}
set.seed(1)
train.index <- sample(row.names(NvBay3), 0.6*dim(NvBay3)[1])  # 60 % of data into training set
valid.index <- setdiff(row.names(NvBay3), train.index)  # 40 %  into validation set
train.df <- NvBay3[train.index, ] # assigning the train.index into data frame
valid.df <- NvBay3[valid.index, ] # assigning the validation index into data frame
train <- NvBay3[train.index, ] # Making a copy of the data frame train.df
valid = NvBay3[train.index,]#  Making a copy of the data frame valid.df
```

A. Create a pivot table for the training data with Online as a column variable, CC as a row variable, and Loan as a secondary row variable. The values inside the table should convey the count. In R use functions melt() and cast(), or function table().

Pivot table For CreditCard , Personal loan as row variables and Online in column.

```{r}
library(reshape2)
melt = melt(train,id=c("CreditCard","Personal.Loan"),variable= "Online")
cast=dcast(melt,CreditCard+Personal.Loan~Online)  # dcast is to convert the data in CC , Personal loan and Online into pivot table.
cast[,c(1,2,3,14)] # casting column no 14 which credit card  and 1 , 2 , 3 column is  , personal loan, ID

```

B. Consider the task of classifying a customer who owns a bank credit card and is actively using online banking services. Looking at the pivot table, what is the probability that this customer will accept the loan offer? [This is the probability of loan acceptance (Loan = 1) conditional on having a bank credit card (CC = 1) and being an active user of online banking services (Online = 1)].
```{r}

Loan_CC1 <- 77/3000  # 77 is the value for Loan and CC =1 as per pivot table.  and 3000 is the total count of 1924 + 198 + 801 + 77 = 3000
Loan_CC1 # which is 26 %.
```


C. Create two separate pivot tables for the training data. One will have Loan (rows) as a function of Online (columns) and the other will have Loan (rows) as a function of CC.
```{r}

melt_1 = melt(train,id=c("Personal.Loan"),variable = "Online") # Melting Personal loan and Online data into pivot table

melt_2 = melt(train,id=c("CreditCard"),variable = "Online") # Melting Credicard data with reference to online

cast_1 =dcast(melt_1,Personal.Loan~Online) # Casting Personal loan and online values
cast_2=dcast(melt_2,CreditCard~Online) # Casting Personal loan and online values
Loanonline=cast_1[,c(1,13)]
LoanCC = cast_2[,c(1,14)]

Loanonline # indicates  personal loan count in reference with online
LoanCC # Indicates Credit Card count in reference with online.
```

D. Compute the following quantities [P (A | B) means “the probability of A given B”]:
P (CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors)
P(Online=1|Loan=1)
P (Loan = 1) (the proportion of loan acceptors)
P(CC=1|Loan=0)
P(Online=1|Loan=0)
P(Loan=0)

```{r}

table(train[,c(14,10)])  # Creating a pivot table for column 14 and 10 which is credit card and personal loan from training

table(train[,c(13,10)])  #  Creating a pivot table for column 13 and 10 which is online and personal loan from training

table(train[,c(10)]) # Pivot table for Personal loan.  There are 2725 and 275 from training
```
P (CC = 1 | Loan = 1)
```{r}

CCLoan_1 = 77/(77+198) # by referring the above pivot table we can get the CC= 1 and Loan = 1 values, which is 77 divided by  CC = 0 and PL 1 198
CCLoan_1
```
P(Online=1|Loan=1)
```{r}
ONLoan_1 =166/(166+109) # by referring the above pivot table we can get the online = 1 and Loan = 1 values
ONLoan_1
```
P (Loan = 1)
```{r}
Loan_1 =275/(275+2725) # by referring the above pivot table we can get the Loan = 1
Loan_1
```
P(CC=1|Loan=0)
```{r}
CCLoan_01= 801/(801+1924) # by referring the above pivot table we can get the CC = 1 and Loan = 0 values / CC 0 and PL 0
CCLoan_01
```

P(Online=1|Loan=0)
```{r}
ON1L0= 1588/(1588+1137)  # by referring the above pivot table we can get the online = 1 and Loan = 0 values
ON1L0
```
P(Loan=0)

```{r}

Loan_0= 2725/(2725+275)  # by referring the above pivot table we can get the Loan = 0 values
Loan_0
```


E. Use the quantities computed above to compute the naive Ba1 probability P(Loan = 1 | CC = 1, Online = 1).
```{r}
Naivebayes = ((77/(77+198))*(166/(166+109))*(275/(275+2725)))/(((77/(77+198))*(166/(166+109))*(275/(275+2725)))+((801/(801+1924))*(1588/(1588+1137))*2725/(2725+275)))

Naivebayes # 90 % is the probability
```



F. Compare this value with the one obtained from the pivot table in (b). Which is a more accurate estimate? 9.05% are very similar to the 9.7% the difference between the exact method and the naive-baise method is the exact method would need the the exact same independent variable classifications to predict, where the naive bayes method does not.


```{r}
library(caret)
library(e1071)
naive.train1 = train.df[,c(10,13,14)] # training data is from Personal loan, Creditcard and online. column
naive.test1 =valid.df[,c(10,13,14)] # testing set data from the same columns of data
naivebayes = naiveBayes(Personal.Loan~.,data=naive.train1) # applying naivebayes algorithm to personal loan and training data.
naivebayes

```


G. Which of the entries in this table are needed for computing P (Loan = 1 | CC = 1, Online = 1)? In R, run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P (Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E).

Answer:

For the naivebayes, it is same output that we have got in the manual calculation prcoess. (.280)(.603)(.09)/(.280.603.09+.29.58.908) = .09 which is the same as the manual calculation process.

